---
output: html_document
---

```{r setup, include = FALSE, warning = F, message = F, echo = F}
knitr::opts_chunk$set(echo = TRUE, message = F, echo = F, warning = F)

library(tidyverse)
library(lubridate)
library(WtRegDO)
library(here)
library(GGally)

# load results from script files
load(file = here('data', 'outputpar.RData'))
load(file = here('data', 'outputkinst.RData'))
load(file = here('data', 'outputmetdt.RData'))
load(file = here('data', 'outputmetptdt.RData'))

fls <- c('outputpar', 'outputkinst', 'outputmetdt', 'outputmetptdt')
names(fls) <- c('Bayes original', 'Bayes K inst', 'Bayes K inst, Met day', 'Bayes K inst, Met day, p/t est.')

bsmetab <- NULL
for(fl in fls){
  
  dat <- get(fl)
  
  out <- dat %>% 
    select(
      Pg = GPP, 
      Rt = ER, 
      NEM = NEP,
      Date
    ) %>% 
    mutate(
      Rt = -1 * Rt,
      Date = ymd(Date) 
    ) %>% 
    gather('var' ,'val', -Date) %>% 
    mutate(typ = names(fls)[fl == fls])
  
 bsmetab <- bind_rows(bsmetab, out)
 
}

load(file = here('data/DO_APNERR2012_6_12_0.8.RData'))
tmp <- DO_APNERR2012_6_12_0.8 %>%
  select(DateTimeStamp, DO_mgl = DO_nrm, Temp, Sal, ATemp, BP, WSpd, totpar, Tide)

opmetab <- ecometab(tmp, tz = 'America/Jamaica', lat = 29.75, long = -85, gasex = 'Wanninkhof', gasave = 'instant', metab_units = 'grams') %>% 
  gather('var', 'val', -Date) %>% 
  mutate(typ = 'Odum')

toplo <- bind_rows(bsmetab, opmetab) %>% 
  mutate(
    typ = factor(typ, levels = c(names(fls), 'Odum'))
  )

toplo2 <- toplo %>% 
  spread(typ, val) %>% 
  select(-Date)

toplo2nem <- toplo2 %>% 
  filter(var == 'NEM') %>% 
  select(-var)
toplo2pg <- toplo2 %>% 
  filter(var == 'Pg') %>% 
  select(-var)
toplo2rt <- toplo2 %>% 
  filter(var == 'Rt') %>% 
  select(-var)
```

# Comparing Bayes and Odum estimates {.tabset}

Evaluated 2012 Apalachicola NERRS data, detided DO Time series using 6, 12, 0.8 window widths.  Gas exchange estimated with Wanninkhof. 

* __Bayes original__ Stripped down BASEmetab code
* __Bayes K inst__ Stripped down BASEmetab code, instantaneous K
* __Bayes K inst, Met day__ Stripped down BASEmetab code with instantaneous K and metabolic day
* __Bayes K inst, Met day, p/t est.__ Stripped down BASEmetab code with instantaneous K, metabolic day, p and theta parameters estimated
* __Odum__ Odum method from WtRegDO, instantaneous K and metabolic day

## Overall plots

```{r, fig.height = 7, fig.width = 7}
p1 <- ggplot(toplo, aes(x = Date, y = val, colour = var, fill = var)) + 
  geom_line() + 
  geom_point() + 
  theme_minimal() + 
  facet_wrap(~typ, ncol = 1) + 
  theme(
    strip.background = element_blank()
  ) +
  labs(
    y = expression(paste('g ', O [2], ' ', m^-2, d^-1))
  ) + 
  theme(
    axis.title.x = element_blank(), 
    legend.title = element_blank()
  )
p1
```

## NEM 

```{r, fig.height = 7, fig.width = 7}
lowerfun <- function(data, mapping, lims = c(-2, 1)){
  ggplot(data = data, mapping = mapping)+
    geom_point() + 
    geom_abline(slope = 1, intercept = 0, color = 'blue') + 
    scale_x_continuous(limits = lims) + 
    scale_y_continuous(limits = lims)
}  

ggpairs(toplo2nem, lower = list(continuous = wrap(lowerfun))) + 
  theme_minimal() + 
  labs(
    title = "NEM"
  )
```

## Pg

```{r, fig.height = 7, fig.width = 7}
lowerfun <- function(data, mapping, lims = c(-1, 2.5)){
  ggplot(data = data, mapping = mapping)+
    geom_point() + 
    geom_abline(slope = 1, intercept = 0, color = 'blue') + 
    scale_x_continuous(limits = lims) +
    scale_y_continuous(limits = lims)
}  

ggpairs(toplo2pg, lower = list(continuous = wrap(lowerfun))) + 
  theme_minimal() + 
  labs(
    title = "Pg"
  )
```

## Rt

```{r, fig.height = 7, fig.width = 7}
lowerfun <- function(data, mapping, lims = c(-3.5, 1)){
  ggplot(data = data, mapping = mapping)+
    geom_point() + 
    geom_abline(slope = 1, intercept = 0, color = 'blue') + 
    scale_x_continuous(limits = lims) +
    scale_y_continuous(limits = lims)
}  

ggpairs(toplo2rt, lower = list(continuous = wrap(lowerfun))) + 
  theme_minimal() + 
  labs(
    title = "Rt"
  )
```